---
title: "classes Spatial Trend Analysis"
output:
  pdf_document: default
  html_document: default
---
##Packages 
```{r}
library(terra)
library(hydroTSM)
library(trend)



```
##Trend analysis 
###Data
```{r}
chirps<-"C:/Users/jmviq/Desktop/R_course_and/L2_Spatiotemporal_Statistics/CHIRPS_annual"
shape<-"C:/Users/jmviq/Desktop/R_course_and/L2_Spatiotemporal_Statistics/Countries/"
chirps<-list.files(chirps,full.name=TRUE)
chirps<-rast(chirps)
shape<-vect(shape)
##Ethiopia 
shape<-shape[7,]
plot(shape)
```

###Crop
```{r}
chirps<-crop(chirps,shape,snap="out")
chirps<-mask(chirps,shape)
plot(chirps)
```
###MAN-KENDALL TREND
####non parametric rank-based statistical technique /n
####null=no monotonic trend /n
####trend can be positive, negative or non-null /n
####probability of 5%
####Sen slope=magnitude by years 

##Aplying the kendall test
###Funtion for KT
```{r}
get_pval<-function(x){
  if(!any(is.na(x))){
   mk<-mk.test(x)
   pval<-mk$p.value
  } else {
     pval<-NA
  }
    return(pval)
  }

```
##Applying the funtion 
```{r}
pval_chirps<-app(chirps,get_pval)
plot(pval_chirps)
hist(pval_chirps)
print(pval_chirps)
```
##Slope(betha/t)
####Funtion for SEN slope
```{r}
get_slope<-function(x, conf.level=0.95,n.years=nlyr(chirps)){
  if(!any(is.na(x))){
  ss<-sens.slope(x,conf.level = conf.level)
  slope<-ss$estimates*n.years
  }else{
    slope<-NA
    
  }
  return(slope)
  
}
```

##Applying the funtion slope  
```{r}
svals_chirps<-app(chirps,get_slope)
plot(svals_chirps)
lines(shape)
hist(svals_chirps)
print(svals_chirps)
```
##Masking changes
```{r}
sig_changes<-pval_chirps
sig_changes[sig_changes>0.05]<-NA
sig_changes[sig_changes>0]<-1
```
##Adding direction 
```{r}
unit_changes<-svals_chirps/abs(svals_chirps)
sig_changes<-unit_changes*sig_changes
plot(sig_changes)
lines(shape)


```

##Results 
```{r}
sig_changes<-abs(sig_changes)
svals_chirps<-svals_chirps*sig_changes
plot(svals_chirps)
lines(shape)


```